%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% chapter4.tex
%% NOVA thesis document file
%%
%% Chapter with lots of dummy text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Conceptual Solution}
\label{cha:conceptual_solution}

% Conceptual Solution
% Introduction
The comprehensive examination of the two fundamental pillars that constitute \gls{sdvn} and \gls{sdvn} itself has provided insights into the potential usefulness of this technology in real-world scenarios, thereby enabling the identification of an ideal practical deployment for this technology. As such, in this section, this ideal view created from the accumulated knowledge presented in this document will be described.
Before proceeding to the specifics, it is important to note that this perspective describes some hypothetical optimal objectives for an implementation. Consequently, any practical implementations developed in this thesis will be constrained by the availability and costs of the requisite hardware and the availability and compatibility of the related software, in addition to timing and implementation constraints.    

% Function
\section{The impact of interoperability}
In light of the requirement for interoperability, as delineated in section \ref{subsub:mandated_interoperability}, an \gls{sdn}-compatible \gls{obu} deployment is only viable in real-world scenarios if it is capable of communicating with \gls{etsi} conformant devices. This limitation for implementation is only applicable in real-world deployments, as the devices must comply with, or at the very least be compatible and able to collaborate with, the surrounding protocols in order to be effective. No existing software stack for \gls{sdn} is compatible with the \gls{its}-G5 protocol stack. Consequently, the most straightforward approach to create a compatible environment is to utilize P4’s capacity to reprogram the packet logic of a network. 
The vast majority of research papers on \gls{sdvn} only implement the first generation of \gls{sdn}, with the few that delve into P4 and its advantages only doing so on a theoretical level, never providing an implementation platform. To the best of our knowledge, and in accordance with the findings of Sarpong et al.\cite{sarpong_potential_2023}, no work has been published using P4 as the data plane technology in a \gls{sdvn} implementation. 
It is important to acknowledge that the aforementioned limitation does diminish the freedom introduced by P4. However, it does not entirely negate its advantages, because the ability to modify network protocols enables protocol experimentation, thereby allowing existing protocols to be tested, validated, and optimized with greater ease.
These observations lead us to conclude that the optimal \gls{sdvn} scenario would entail the use of a P4 device that emulates the protocol fields of the \gls{etsi}-defined protocol stack, modified only to accommodate changes to the controller. The aforementioned conclusion, derived from the collected theoretical evidence, suggests that OpenFlow-based solutions are not a viable option. Nevertheless, in order to corroborate this hypothesis, the requisite practical test will be conducted.
% Controller placement
\section{Controller placement}
The controller brings us to the second issue of controller distribution. Section \ref{subsec:control_plane} displays the three different possible locations where the controller can be implemented, but most \gls{sdn} developments only implement controller instances in the wired infrastructure. Due to the issues touched upon in subsection \ref{subsub:mandated_interoperability} and \ref{subsec:issues_with_maintaining_a_global_network_view} devices must be able to function in tandem with traditional devices and without connection to infrastructure. 
Ku et al.\cite{ku_towards_2014}, although the first, had already raised these issues deserved attention by incorporating recovery mechanisms in their experimental scenario. They used a backup traditional algorithm when the device lost connection with the controller, demonstrating that \gls{sdvn} must have failure recovery mechanisms to ensure the network works in all scenarios.
The present research will not be able to examine the particulars of the controller, such as identifying the optimal controller distribution scheme in software-defined vehicular environments. This is because it is essential to examine all subsystems within the \gls{its} domain in order to achieve meaningful results, and this thesis will focus on the \gls{its} vehicle subsystem.
However, for the aforementioned reasons, it should be considered fundamental for a practical \gls{sdvn} deployment to contain a local controller able to operate each individual device and take over with something similar to traditional software if the connection to the infrastructure is lost. 
    